Supplementary Material for _A Guide to Pre-processing High-Frequency Animal Tracking Data_

### 2021-05-20 supplement/05_calibration_data.Rmd
check for remotes
new figure output paths

```{r install_atlastools, eval=TRUE}
install.packages("remotes")

@@

# save figure
ggsave(fig_data_raw,
  filename = "figures/fig_calibration_raw.png",
  width = 185 / 25
)
```

![The raw data from a calibration exercise conducted around the island of Griend in the Dutch Wadden Sea. A handheld WATLAS tag was used to examine how ATLAS data compared to GPS tracks, and we use the WATLAS data here to demonstrate the basics of the pre-processing pipeline, as well as validate the residence patch method. It is immediately clear from the figure that the track shows location errors, both in the form of point outliers as well as small-scale errors around the true location.](figures/fig_calibration_raw.png)

@@

# save result
ggsave(fig_data_bbox,
  filename = "figures/fig_calib_bbox.png",
  width = 185 / 25
)
```

![Removal of a point outlier using the function `atl_filter_bounds`. The point outlier (black point) is removed based on its X coordinate value, with the data filtered to exclude positions with an X coordinate < 645,000 in the UTM 31N coordinate system. Positions that are retained are shown in green.](figures/fig_calib_bbox.png)

@@

### Get 95th percentile of speed and angle

@@

```{r echo=FALSE}
# plot filtered data
fig_speed_outliers <-
  ggplot() +
  geom_point(
    data = data,
    aes(x, y,
      col = speed_in < 25.6
    ),
    size = 0.3, alpha = 0.5,
    show.legend = F
  ) +
  annotate(
    geom = "rect",
    xmin = 650785, xmax = 653250,
    ymin = 5904450, ymax = 5906133,
    fill = NA, col = "grey20"
  ) +
  annotate(
    geom = "text",
    x = 653250,
    y = 5906133 + 500,
    label = "(b)",
    # size = 10,
    fontface = "bold"
  ) +
  annotate(
    geom = "rect",
    ymin = 5901650, ymax = 5903100,
    xmin = 650000, xmax = 650600,
    fill = NA, col = "grey20"
  ) +
  annotate(
    geom = "text",
    y = 5901650 - 300,
    x = 650600,
    label = "(c)",
    # size = 10,
    fontface = "bold"
  ) +
  scale_color_manual(values = c("grey", pal[3])) +
  ggthemes::theme_few() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank()
  ) +
  ggspatial::annotation_scale(location = "tl") +
  coord_sf(crs = 32631)
# save
ggsave(fig_speed_outliers,
  filename = "figures/fig_speed_outlier.png",
  width = 170 / 25, height = 170 / 25
)
```

@@

![Improving data quality by filtering out positions that would require unrealistic movement. We removed positions with speeds $\geq$ 15 m/s, which is the fastest possible speed in this calibration data, part of which was collected in a moving boat around Griend. Grey positions are removed, while green positions are retained. Rectangles indicate areas expanded for visualisation in following figures.](figures/fig_speed_outlier.png)

@@

ggsave(fig_smooth,
  filename = "figures/fig_calib_median_smooth.png",
  width = 90 / 25, height = 90 / 25
)
```

![Reducing small-scale location error using a median smooth with a moving window $K$ = 5. Median smoothed positions are shown in green, while raw, unfiltered data is shown in grey. Median smoothing successfully recovers the likely path of the track without a loss of data. The area shown is the upper rectangle from Fig. 1.3.](figures/fig_calib_median_smooth.png)

@@

ggsave(fig_smooth_thin,
  filename = "figures/fig_calib_smooth_thin.png",
  width = 90 / 25, height = 90 / 25, dpi = 300
)
```

@@

# # save combined figure
ggsave(figure_walkthrough,
  filename = "figures/fig_walkthrough.png",
  height = 90, width = 170, units = "mm"
)
```

![Thinning by aggregation over a 30 second interval (down from 1 second) preserves track structure while reducing the data volume for computation. Here, thinned positions are shown as purple squares, with the size of the square indicating the number of positions within the 30 second bin used to obtain the average position. Green points show the median smoothed data from Fig. 1.4, while the raw data are shown in grey. The area shown is the upper rectangle in Fig. 1.3.](figures/fig_calib_smooth_thin.png)

>>>>

```{r install_atlastools, eval=TRUE}
if (!require(remotes)) {
  install.packages("remotes", repos = "http://cran.us.r-project.org")
}

@@

# save figure
ggsave(fig_data_raw,
  filename = "supplement/figures/fig_calibration_raw.png",
  width = 185 / 25
)
```

![The raw data from a calibration exercise conducted around the island of Griend in the Dutch Wadden Sea. A handheld WATLAS tag was used to examine how ATLAS data compared to GPS tracks, and we use the WATLAS data here to demonstrate the basics of the pre-processing pipeline, as well as validate the residence patch method. It is immediately clear from the figure that the track shows location errors, both in the form of point outliers as well as small-scale errors around the true location.](supplement/figures/fig_calibration_raw.png)

@@

# save result
ggsave(fig_data_bbox,
  filename = "supplement/figures/fig_calib_bbox.png",
  width = 185 / 25
)
```

![Removal of a point outlier using the function `atl_filter_bounds`. The point outlier (black point) is removed based on its X coordinate value, with the data filtered to exclude positions with an X coordinate < 645,000 in the UTM 31N coordinate system. Positions that are retained are shown in green.](supplement/figures/fig_calib_bbox.png)

@@

### Get 90th percentile of speed and angle

@@

```{r echo=FALSE}
# plot filtered data
fig_speed_outliers <-
  ggplot() +
  geom_point(
    data = data,
    aes(x, y,
      col = speed_in < 25.6
    ),
    size = 0.3, alpha = 0.5,
    show.legend = F
  ) +
  annotate(
    geom = "rect",
    xmin = 650785, xmax = 653250,
    ymin = 5904450, ymax = 5906133,
    fill = NA, col = "grey20"
  ) +
  annotate(
    geom = "text",
    x = 653250,
    y = 5906133 + 500,
    label = "(b)",
    # size = 10,
    fontface = "bold"
  ) +
  annotate(
    geom = "rect",
    ymin = 5901650, ymax = 5903100,
    xmin = 650000, xmax = 650600,
    fill = NA, col = "grey20"
  ) +
  annotate(
    geom = "text",
    y = 5901650 - 300,
    x = 650600,
    label = "(c)",
    # size = 10,
    fontface = "bold"
  ) +
  scale_color_manual(values = c("grey", pal[3])) +
  ggthemes::theme_few() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank()
  ) +
  ggspatial::annotation_scale(location = "tl") +
  coord_sf(crs = 32631)

# save
ggsave(fig_speed_outliers,
  filename = "supplement/figures/fig_speed_outlier.png",
  width = 170 / 25, height = 170 / 25
)
```

@@

![Improving data quality by filtering out positions that would require unrealistic movement. We removed positions with speeds $\geq$ 15 m/s, which is the fastest possible speed in this calibration data, part of which was collected in a moving boat around Griend. Grey positions are removed, while green positions are retained. Rectangles indicate areas expanded for visualisation in following figures.](supplement/figures/fig_speed_outlier.png)

@@

ggsave(fig_smooth,
  filename = "supplement/figures/fig_calib_median_smooth.png",
  width = 90 / 25, height = 90 / 25
)
```

![Reducing small-scale location error using a median smooth with a moving window $K$ = 5. Median smoothed positions are shown in green, while raw, unfiltered data is shown in grey. Median smoothing successfully recovers the likely path of the track without a loss of data. The area shown is the upper rectangle from Fig. 1.3.](supplement/figures/fig_calib_median_smooth.png)

@@

ggsave(fig_smooth_thin,
  filename = "supplement/figures/fig_calib_smooth_thin.png",
  width = 90 / 25, height = 90 / 25, dpi = 300
)
```

@@

# # save combined figure
ggsave(figure_walkthrough,
  filename = "supplement/figures/fig_walkthrough.png",
  height = 90, width = 170, units = "mm"
)
```

![Thinning by aggregation over a 30 second interval (down from 1 second) preserves track structure while reducing the data volume for computation. Here, thinned positions are shown as purple squares, with the size of the square indicating the number of positions within the 30 second bin used to obtain the average position. Green points show the median smoothed data from Fig. 1.4, while the raw data are shown in grey. The area shown is the upper rectangle in Fig. 1.3.](supplement/figures/fig_calib_smooth_thin.png)
